---
title: "Tournament Variance"
author: "Lee Przybylski"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(reactable)
```

## Background

We will simulate a tournament of 12 teams in 2 formats.  The first is a double
elimination format with games that are best of 3 to 11 on the main draw and 1 
game to 15 in the back draw.  The second is a round robin with 2 pools of 6 
teams.  In pool play, every team plays the other 5 teams in 1 game to 11.  The 
top 3 from each pool advance to a 6 team playoff consisting of games to 15.  The
medal matches are best of 3 to 11.  We assume the seeding is perfect according
to the probability of each team winning a rally.

## Functions

First we need some utility functions.

```{r}
#A logistic function to get binomial probability team 1 wins
logistic <- function(a1 = 0, a2 = 0){
 p <- 1/(1+ exp(a2 - a1))
 return(p)
}

play_rally <- function(r1 = 0, r2 = 0, n = 1){
 p <- logistic(r1, r2)
 winner <- ifelse(rbinom(n, 1, p) > 0, 1, 2)
 return(winner)
}

play_singles_game <- function(
  a_serve = c(0, 0),
  a_return = c(a_serve[1] + 0.01, a_serve[2] + 0.01),
  f_score = 11, 
  server = NA
  ){
 if(is.na(server)) server <- sample(1:2, size = 1)
 scores <- c(0,0)
 a <- matrix(c(a_serve, a_return), nrow = 2, ncol = 2)
 while(max(scores) < 11 | abs(scores[1] - scores[2]) < 2){
  #specify rally win probs
  a1 <- a[1, ifelse(server == 1, 1, 2)]
  a2 <- a[2, ifelse(server == 2, 1, 2)]
  #play the rally
  rally_outcome <- play_rally(a1, a2)
  #assign the point and server for the next rally
  if(rally_outcome == server){
    scores[rally_outcome] <- scores[rally_outcome] + 1
  }else{
   server <- rally_outcome
  }
  
 }
 return(scores)
}

play_singles_match <- function(
  a_serve = c(0, 0), 
  a_return = c(a_serve[1] + 0.01, a_serve[2] + 0.01), 
  g_max = 3, 
  f_score = 11){
 assertthat::assert_that(g_max %% 2 == 1, 
                         msg = "check g_max; invalid number of games"
                         )
 #Choose who serves first
 server <- sample(1:2, 1)
 #initialize games and scores
 games <- c(0,0)
 scores <- ""
 while(max(games) < (g_max + 1)/2){
  game_i <- play_singles_game(a_serve, a_return, f_score, server = server)
  winner <- which.max(game_i)
  games[winner] <- games[winner] + 1
  scores <- paste(scores, paste(game_i[1], game_i[2], sep = "-"), " ", sep = "")
  server <- ifelse(server == 1, 2, 1)
 }
 
 winner <- which.max(games)
 #Drop the last space from scores
 #scores <-
 
 output <- data.frame(winner = winner, 
                      scores = scores, 
                      games1 = games[1], 
                      games2 = games[2],
                      a1serve = a_serve[1],
                      a1return = a_return[1],
                      a2serve = a_serve[2],
                      a2return = a_return[2])
 return(output)
}
```

Next, we need functions to play the bracket.  With 12 team double elimination, 
we give the top 4 seeds a bye.

```{r}
#field is a df with listing the players and their power ratings
field <- data.frame(rating = rnorm(12, sd = 0.2)) %>% 
 as_tibble() %>% 
 arrange(-rating) %>% 
 mutate(seed = 1:12, .before = everything())

play_de16_bracket <- function(field){
 #List the matches
 rd1_matches <- data.frame(match = rep(1:8, each = 2), 
                           seed = c(1, 16,
                                    8, 9,
                                    5, 12,
                                    4, 13,
                                    3, 14,
                                    6, 11,
                                    7, 10,
                                    2, 15
                                    )
                           )
}
```